\Large{Adversarial Training against Systematic Uncertainty}

### 2016-09-27 tex/paper.tex
save

\begin{figure*}
    \label{fig:toy}
    \begin{center}
        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
        \vspace{-0.8cm}
        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{Toy example.
    (Top left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
    (Top right) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ closer to $5$,
       hence explaining the dependency.
    (Bottom left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now almost identical to each other, indicating only a
       small dependency on $Z$.
    (Bottom right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function to erase the dependency on $Z$.
    }
\end{figure*}

\begin{figure}
    \label{fig:toy-training}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
\end{figure}

As a guiding toy example, let us consider the binary classification problem of 2D
data drawn from multivariate gaussians with equal priors, such that
\begin{align}
    x &\sim {\cal N}(\mu=(0,0), \sigma=2 \times \mathbb{1}) \text{ when } Y=0, \\
    x &\sim {\cal N}(\mu=(1,-1+z), \sigma=2 \times \mathbb{1}) \text{ when } Y=1.
\end{align}
The nuisance parameter $Z$ represents in this case our
uncertainty about the exact location of the mean of the second gaussian. Our goal is to
build a classifier $f(\cdot;\theta_f)$ for predicting $Y$ given $X$, but such that
the probability distribution of $f(X;\theta_f)$ is invariant with respect to the
nuisance parameter $Z$.

Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we train
a classifier $f$ minimizing $L_f(\theta_f)$ without any adversary.
\glnote{to be continued}

>>>>

\begin{figure*}
    \label{fig:toy}
    \begin{center}
        \includegraphics[width=0.48\textwidth]{figures/f-plain.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-plain.pdf}\\
        \vspace{-0.8cm}
        \includegraphics[width=0.48\textwidth]{figures/f-adversary.pdf}
        \includegraphics[width=0.48\textwidth]{figures/surface-adversary.pdf}
    \end{center}
    \vspace{-1cm}
    \caption{Toy example.
    (Top left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$
       when $f$ is built without adversarial training. The resulting densities
       are clearly dependent on $Z$, indicating that $f$ is not pivotal.
    (Top right) The corresponding decision surface, highlighting
       the fact that samples are easier to classify for values of $Z$ closer to $5$,
       hence explaining the dependency.
    (Bottom left) Probability densities of the decision scores for $Z=0$, $2.5$ and $5$ when $f$ is
       built with adversarial training, as outlined in Section~\ref{sec:method}.
       The resulting densities are now closer to each other, indicating only a
       small dependency on $Z$.
    (Bottom right) The corresponding decision surface, illustrating how adversarial
       training bends the decision function to erase the dependency on $Z$.
    }
\end{figure*}

\begin{figure}
    \label{fig:toy-training}
    \includegraphics[width=0.48\textwidth]{figures/training.pdf}
    \vspace{-1cm}
    \caption{Toy example. Training curves for $L_f(\theta_f)$, $L_r(\theta_f, \theta_r)$ and $L_f(\theta_f) - \lambda L_r(\theta_f, \theta_r)$.}
\end{figure}

As a guiding toy example, let us consider the binary classification problem of 2D
data drawn from multivariate gaussians with equal priors, such that
\begin{align}
    x &\sim {\cal N}(\mu=(0,0), \sigma=2 \times \mathbb{1}) \text{ when } Y=0, \\
    x &\sim {\cal N}(\mu=(1,-1+z), \sigma=2 \times \mathbb{1}) \text{ when } Y=1.
\end{align}
The continuous nuisance parameter $Z$ represents in this case our
uncertainty about the exact location of the mean of the second gaussian. Our goal is to
build a classifier $f(\cdot;\theta_f)$ for predicting $Y$ given $X$, but such that
the probability distribution of $f(X;\theta_f)$ is invariant with respect to the
nuisance parameter $Z$.

Assuming a uniform prior over the interval ${\cal Z} = [0;5]$, we start by
generating training data $\{ x_i, y_i, z_i \}_{i=1}^N$, from which we build
a classifier $f$ minimizing $L_f(\theta_f)$ without adversarial training.
\glnote{to be continued}
